What to simulate? Inferring the right direction for mental rotation

### 2014-02-07 man/cogsci-2014/cogsci-2014.tex
Misc small changes (including Fig.-->Figure and Eq.-->Equation)

\begin{abstract}
  When people use mental imagery, how do they decide \textit{which}
  images to generate? To answer this question, we explored how mental
  simulation should used in the classic psychological task of
  determining if two images depict the same object in different
  orientations \cite{Shepard1971}. Through a rational analysis of
  mental rotation, we formalized four models and compared them to
  human performance. \TODO{None of the models were a particularly
    satisfying explanation of human behavior, leading us to suggest
    that the question of ``what to simulate?'' is more difficult than
    has previously been assumed.} While only exploratory, this work
  provides the foundation necessary to begin answering this question.

@@

Mental rotation provides a simple example of the decision problem
posed by simulation.  In the classic experiment by
\citeA{Shepard1971}, participants viewed images of three-dimensional
objects and had to determine whether the images depicted the same
object (which differed by a rotation) or two separate objects (which
differed by a reflection and a rotation). They found that people's
response times had a strong linear correlation with the minimum angle
of rotation, a result which led to the conclusion that people solve
this task by ``mentally rotating'' the objects until they are
congruent.  This explanation leaves several questions unanswered,
however. How do people know the axis around which to rotate the
objects? If the axis is known, how do people know which direction to
rotate the objects?  And finally, how do people know how long to
rotate?

@@

We define $p(\Xb\ \vert\ \Xa, \theta, h)$ to be the similarity between
$\Xb$ and a transformation of $\Xa$: $p(\Xb\ \vert\ \Xa, \theta, h):=
S(\Xb, f(\Xa, h, \theta))$.  We do not know which vertices of $\Xb$
correspond to which vertices of $\Xa$, so $S$ must marginalize over
the set of possible mappings. For brevity, let
$\mathbf{X}_m=\M\cdot{}f(\Xa, h, \theta)$ where $\M$ is a permutation
matrix. Then:
\begin{equation}
  S(\Xb, f(\Xa, h, \theta)):=\frac{1}{2N} \sum_{\M} \prod_{n=1}^N \mathcal{N}(\mathbf{x}_{bn}\ \vert \ \mathbf{x}_{mn}, \I\sigma_S^2),
  \label{eq:similarity}
\end{equation}
where $2N$ is the total number of possible mappings,\footnote{It is
  $2N$ and not $N^2$ because, in polar coordinates, vertices are
  always connected to their two nearest neighbors in the $\theta$
  dimension.}and $\sigma_S^2=0.15$ is the variance of the
similarity. Example similarity curves are shown in
Fig.~\ref{fig:model-traces}.

@@

To summarize, we approximate the likelihood term of
Eq.~\ref{eq:lh-ratio} using the similarity function defined in
Eq.~\ref{eq:similarity}. Because we assume mental rotations are
performed sequentially, this similarity can only be computed for the
actions listed in Eq.~\ref{eq:actions}.

@@

In order to approximate Eq.~\ref{eq:lh-ratio} using samples of the
similarity function, we must decide \textit{which} places to sample
and \textit{when} stop sampling. The models below differ in how they
make these decisions (see Fig.~\ref{fig:model-traces} for examples of
their behavior).

@@

To determine the correct rotation, we solve for the rotation matrix by
computing $(\Xa \F_h^T)_\mathrm{left}^{-1}\cdot{}\Xb$, where
$(\Xa\F_h^T)_\mathrm{left}^{-1}$ is the left inverse of
$\Xa\F_h^T$. We then check each $h$ to see if the computation produces
a valid rotation matrix; the $h$ that does is the correct
hypothesis. This approach gives us the true value of $\theta$, so
Eq.~\ref{eq:lh-ratio} becomes a generalized likelihood ratio test,
where $\theta$ is set to the MLE value, rather than being
marginalized:
\begin{equation}
  \ell = \frac{\max_\theta p(\Xb\ \vert\ \Xa, \theta, \hi)p(\hi)}{\max_\theta p(\Xb\ \vert\ \Xa, \theta, \hf)p(\hf)}.
  \label{eq:mle-lh-ratio}
\end{equation}

@@

What if the point correspondences are unknown? An intuitive and simple
model which obeys this constraint first rotate in the direction that
increases similarity, and then stop rotating once a ``match'' is
found, as determined by a threshold on the value of $S$. If all
rotations have been exhausted, then flip, and try rotating again. We
assume that the locations where $S$ is greater than the threshold
correspond to the true $\theta$ (or points near the true
$\theta$). So, as with the \Oc{} model, Eq.~\ref{eq:lh-ratio} becomes
a likelihood ratio.

@@

In the current formulation of the problem, choosing the threshold is
straightforward because we know both the exact geometry of the shapes
and that a linear transformation exists which will perfectly align
them. However, this choice is not always clear \textit{a priori}, as
the global optimum depends on many factors (e.g., shape complexity,
dimensionality, perceptual uncertainty, and whether the shapes are
identical).  One way to deal with the problem of choosing a threshold
would simply be to perform an exhaustive search or use a global
optimization strategy. However, these strategies would not result in
the linear response time found by \citeA{Shepard1971}. A second
alternative, which could potentially demonstrate the linear response
time result, might instead just perform a Hill Climbing (\Hc{}) search
for a \textit{local} maximum. As with the \Oc{} and \Th{} models,
Eq.~\ref{eq:lh-ratio} becomes a generalized likelihood ratio test.

@@

While the previous few models have all focused on \textit{searching}
for the global maximum, the goal is really to just approximate
Eq.~\ref{eq:lh-ratio} while still obeying the constraint of sequential
rotations. Based on evidence that people use ``active sampling''
\cite<e.g.,>{Nelson2007,Gureckis:2012gu}, we hypothesize a model that
actively monitors the usefulness of the rotations it is
performing. Instead of searching for a maximum, we maintain a
probability distribution over our \textit{estimate} of
Eq.~\ref{eq:lh-ratio}, and then sample actions which are expected to
improve that estimate.  This strategy has several benefits. First, it
does not make assumptions about the shape or scale of the similarity
function. Second, by choosing to sample places which are informative,
this method implicitly minimizes the amount of rotation.

Formally, we denote $Z_h$ as our estimate of the likelihood for
hypothesis $h$, and write its distribution as:
\begin{equation}
  p(Z_h) = \int \left[\int S(\Xb, f(\Xa, \theta, h))p(\theta)\dif\theta\right] p(S)\dif S,
\end{equation}
where $S$ is the similarity function, and $p(S)$ is a prior over
similarity functions.  This method of estimating an integral is known
in the machine-learning literature as \textit{Bayesian Quadrature}
\cite{Diaconis:1988uo,OHagan:1991tx,Osborne:2012tm}, or BQ.

@@

Assuming independence, we can now write $p(Z_h)\approx\mathcal{N}(Z_h\
\vert\ \mathbb{E}[Z_h], \mathbb{V}(Z_h))$, which gives us a
distribution over the likelihood ratio in Eq.~\ref{eq:lh-ratio}:
\begin{equation}
p(\ell)\approx\frac{\mathcal{N}(Z_0\ \vert\ \mathbb{E}[Z_0], \mathbb{V}(Z_0))\ p(\hi)}{\mathcal{N}(Z_1\ \vert\ \mathbb{E}[Z_1], \mathbb{V}(Z_1))\ p(\hf)}.
\end{equation}
This distribution cannot easily be calculated; however, we really are
just interested in whether $Z_0>Z_1$ or $Z_1>Z_0$. Thus, rather than
computing $p(\ell)$, we use $Z_D=Z_0-Z_1$ and compute:
$p(Z_D)\propto\mathcal{N}(\mathbb{E}[Z_0] - \mathbb{E}[Z_1],
\mathbb{V}(Z_0) + \mathbb{V}(Z_1))$.

We then sample new observations until we are at least 95\% confident
that $Z_D\neq 0$. In other words, when $p(Z_D<0)<0.025$, we accept
$\hi$, and when $p(Z_D<0)>0.975$, we accept $\hf$. To determine which
points to sample, we can compute the expected variance of $Z_h$ if we
sampled a new observation at $\theta_a$. From \citeA{Osborne:2012tm}:
\begin{align}
\mathbb{E}&[\mathbb{V}(Z_h|\theta_a)]=\mathbb{V}(Z_h) + \mathbb{E}[Z_h] - \\
&\int \mathbb{E}[Z_h|\theta_{a}]^2 \mathcal{N}(\mu_h(\theta_a), \mathrm{Cov}_h(\theta_a, \theta_a))\dif\log S_h(\theta_a).\nonumber
\label{ref:expected-variance}
\end{align}
We compute this expected variance for each of the actions in
Eq.~\ref{eq:actions}, and then choose the action with the lowest
value.

@@

We randomly generated 20 shapes of five or six vertices (e.g.,
Fig.~\ref{fig:model-traces}). For each shape, we computed 20 ``same''
and 20 ``flipped'' stimuli pairs, with 18 rotations ($\theta$) spaced
at $20^\circ$ increments between $0^\circ$ and $360^\circ$ (with
$0^\circ$ and $180^\circ$ repeated twice, in order to gather an equal
number of responses for each angle between $0^\circ$ and
$180^\circ$). ``Same'' pairs were created by rotating $\Xa$ by
$\theta$; ``flipped'' pairs were first reflected $\Xa$ across the
$y$-axis, then rotated by $\theta$.

@@

\begin{figure*}[t]
  \begin{center}
    \includegraphics[width=\textwidth]{../../figures/D/response_time_scatters.pdf}
    \caption{\textbf{Model vs. human RTs.}}
    \label{fig:human-model-scatters}
  \end{center}
\end{figure*}

@@

For each model, we ran 50 samples for each of the 800 experimental
stimuli. The step size parameter ($\sigma_\epsilon$) was fit to human
RTs for each of the models, resulting in $\sigma_\epsilon=0.6$ for the
\Th{} and \Bq{} models and $\sigma_\epsilon=0.1$ for the \Oc{} and
\Hc{} models. We also ran the models under two different priors,
$p(h=0)=0.5$ (the ``equal'' prior) and $p(h=0)=0.55$ (the ``unequal''
prior). This only had an effect on the stopping criteria for the \Bq{}
model.\footnote{The prior also had a small effect on the accuracy of
  the \Hc{} model, because sometimes the local maxima were small
  enough that Eq.~\ref{eq:mle-lh-ratio} became dominated by the
  prior.}

@@

For analyses of RT, confidence intervals around harmonic means were
computed using a bootstrap analysis using 10000 bootstrap samples
(sampled with replacement). RTs were only considered for correct
responses.  We also used a bootstrap analysis of 10000 bootstrap
samples to compute the confidence intervals around both Spearman
($r_s$) and Pearson ($\rho$) correlations.  Unless otherwise
specified, all correlations were computed over 720 datapoints (one for
each stimulus). For analyses of accuracy, confidence intervals were
computed from a binomial proportion with a Jeffrey's beta prior.  To
test if participants or a model was above chance on a particular
stimulus, we used the same binomial proportion, and simply tested
whether $p>0.05$ for the 50\% mark.

@@

The average RT across all stimuli was \ExpTime{} milliseconds.  The
minimum angle of rotation was significantly monotonically correlated
with average per-stimulus RTs, both for ``flipped''
(\ExpThetaTimeCorrFlipped{}) and ``same'' pairs
(\ExpThetaTimeCorrSame{}). While this replicates the general result of
previous experiments \cite<e.g.,>{Shepard1971,Cooper:1975wp}, our
results are not as perfectly linear. However, \citeA{Gardony:2013gn}
found a constant effect for ``flipped'' pairs; an average of their
results and the classic results would produce something similar to
ours.

@@

To gauge between-subject reliability, we randomly split subjects into
two groups, computed the average RT and accuracy per stimulus for each
group, and then computed correlations for these two measures. We
repeated this procedure 100 times, and from those samples computed the
median correlations and 95\% confidence intervals. For RT, the
between-subjects correlation was \ExpTimeCorr{}, and for accuracy, it
was \ExpAccuracyCorr{}.

@@

As we would expect, the number of actions taken by the \Oc{} model was
perfectly correlated with the minimum angle of rotation
(Fig.~\ref{fig:response-time-accuracy}). Unsurprisingly, the \Oc{}
model was the best fit to human RTs, with a correlation of
\OcTimeCorr{} (Fig.~\ref{fig:human-model-scatters}).  However, the
\Oc{} model was 100\% accurate, and therefore the correlation between
minimum angle and accuracy was undefined.

@@

There was an overall monotonic relationship between the minimum angle
of rotation and the number of actions taken by the \Th{} model
(Fig.~\ref{fig:response-time-accuracy}). However, the relationship
between angle and RT for individual shapes was usually not monotonic
(e.g., Fig.~\ref{fig:response-time-stimulus}).  The \Th{} model was
able to explain a moderate amount of the variance in human RTs, with a
correlation of \ThTimeCorr{}
(Fig.~\ref{fig:human-model-scatters}). Like the \Oc{} model, the \Th{}
model had 100\% accuracy, and thus there was no correlation between
minimum angle and accuracy.

As noted above, we fit $\sigma_\epsilon=0.6$ for the \Th{} model. This
has the interesting effect of actually causing the \Th{} to
\textit{over}rotate (see, for example, the ``flipped'' pair trace in
Fig.~\ref{fig:model-traces}). This is because the step size is large
enough that it sometimes misses the global maximum, and must do
another full rotation to find it. However, when the step size is small
enough to always find the maximum, the fit to human data is slightly
worse.

@@

\begin{figure}[t]
  \begin{center}
    \includegraphics[width=0.48\textwidth]{../../figures/D/response_time_stimulus.pdf}
    \caption{\textbf{Typical RT curves for a single shape.}  These
      plots correspond to the shape shown in
      Fig.~\ref{fig:model-traces}. Left: human curves are either
      linear (as with the ``same'' pairs), or linear and then flat (as
      with the ``flipped'' pairs). Middle: the \Th{} model does not
      have a reliably monotonic relationship with rotation. Right: the
      \Bq{} model is roughly linear.}
    \label{fig:response-time-stimulus}
  \end{center}
\end{figure}

The \Hc{} was the only model for which there was no monotonic
relationship between rotation and RT
(Fig.~\ref{fig:response-time-accuracy}). Moreover, the \Hc{} model was
barely above chance (\HcAccuracy{}) and there were \HcNumChance{}
individual stimuli for which it was not above chance. The \Hc{} model
was not a particularly good predictor of human RTs (\HcTimeCorr{}), as
shown in Fig.~\ref{fig:human-model-scatters}. It was a moderately good
predictor of human accuracy (\HcAccuracyCorr{}).

@@

Like the \Oc{} and \Th{} models, there was an overall monotonic
relationship between rotation and the number of steps taken by the
\Bq{} model (Fig.~\ref{fig:response-time-accuracy}). Unlike the \Th{}
model, this relationship existed for individual shapes as well (e.g.,
Fig.~\ref{fig:response-time-stimulus}).  The \Bq{} model explained
variance in human RTs about as well as the \Th{} model
(Fig.~\ref{fig:human-model-scatters}), with a correlation of
\BqTimeCorr{} for the equal prior and \BqpTimeCorr{} for the unequal
prior.

The \Bq{} model was quite accurate overall (equal prior:
(\BqAccuracy{}; unequal prior: \BqpAccuracy{}). With the equal prior,
there was \BqNumChance{} stimulus for which it was not above chance;
with the unequal prior, there were \BqpNumChance{}. The \Bq{} model
explained about the same amount of variance in people's accuracy as
the \Hc{} model (equal prior: \BqAccuracyCorr{}; unequal prior:
\BqpAccuracyCorr{}).

Because the \Bq{} model relies on Eq.~\ref{eq:lh-ratio} for its
stopping criteria (as opposed to just finding a maximum), the prior
$p(h)$ had a fairly significant effect, as shown in
Fig.~\ref{fig:response-time-accuracy}. With just a small bias of
$p(h=0)=0.55$, there was a clear separation in RTs for ``same'' versus
``flipped'' stimuli; this separation is similar to the trend also
observed in human RTs. While the prior also had an effect on accuracy,
this effect did not reflect the trends in human behavior.

@@

The \Oc{} model was the best predictor of human RTs, but is an
unsatisfying explanation of human behavior because it relies on
\textit{a prior} knowledge that people are unlikely to have about
which points on the two shapes correspond. Similarly, while the \Th{}
was next best in terms of correlation, it makes the unsatisfying
assumption that the global maximum is known. Finally, neither of these
models can explain the difference in people's behavior on ``same'' and
``flipped'' stimuli.  The \Bq{} model was nearly as good as the \Th{}
model; in contrast to the other models, the \Bq{} model makes no
assumptions about people's \textit{a prior} knowledge. Furthermore,
the \Bq{} model's adaptive stopping rule is sensitive to the prior,
and thus provides a possible explanation for why people are slower to
respond on ``flipped'' stimulus pairs.

Additionally, a closer look at performance on individual shapes
reveals several interesting points
(Fig.~\ref{fig:response-time-stimulus} shows curves for one particular
shape; these curves are fairly typical relative to other
shapes). First, the curves for the human RTs on specific shapes tend
to either be linear, or linear first, then flat. Second, while the
\Th{} model appears to capture the concavity of human RTs overall, it
is actually fairly inconsistent within shapes. Third, the linearity of
the \Bq{} model's RT curves \textit{is} consistent within shapes,
though it does not exhibit the same ``plateau'' effect that sometimes
appears in the human curves.

@@

To conclude, we hypothesize that people decide \textit{what} to
simulate through an active sampling strategy similar to that used by
the \Bq{} model. While more research is required to confirm or
disprove this claim, the work presented in this paper provides the
foundation necessary to do so.

>>>>

\begin{abstract}
  When people use mental imagery, how do they decide \textit{which}
  images to generate? To answer this question, we explored how mental
  simulation should be used in the classic psychological task of
  determining if two images depict the same object in different
  orientations \cite{Shepard1971}. Through a rational analysis of
  mental rotation, we formalized four models and compared them to
  human performance. \TODO{None of the models were a particularly
    satisfying explanation of human behavior, leading us to suggest
    that the question of ``what to simulate?'' is more difficult than
    has previously been assumed. While only exploratory, this work
    provides the foundation necessary to begin answering this
    question.}

@@

Mental rotation provides a simple example of the decision problem
posed by simulation.  In the classic experiment by
\citeA{Shepard1971}, participants viewed images of three-dimensional
objects and had to determine whether the images depicted the same
object (which differed by a rotation) or two separate objects (which
differed by a reflection and a rotation). They found that people's
response times had a strong linear correlation with the minimum angle
of rotation, a result which led to the conclusion that people solve
this task by ``mentally rotating'' the objects until they are
congruent.  However, this explanation leaves several questions
unanswered. How do people know the axis around which to rotate the
objects? If the axis is known, how do people know which direction to
rotate the objects?  And finally, how do people know how long to
rotate?

@@

We define $p(\Xb\ \vert\ \Xa, \theta, h)$ to be the similarity between
$\Xb$ and a transformation of $\Xa$: $p(\Xb\ \vert\ \Xa, \theta, h):=
S(\Xb, f(\Xa, h, \theta))$.  We do not know which vertices of $\Xb$
correspond to which vertices of $\Xa$, so $S$ must marginalize over
the set of possible mappings. For brevity, let
$\mathbf{X}_m=\M\cdot{}f(\Xa, h, \theta)$ where $\M$ is a permutation
matrix. Then:
\begin{equation}
  S(\Xb, f(\Xa, h, \theta)):=\frac{1}{2N} \sum_{\M} \prod_{n=1}^N \mathcal{N}(\mathbf{x}_{bn}\ \vert \ \mathbf{x}_{mn}, \I\sigma_S^2),
  \label{eq:similarity}
\end{equation}
where $2N$ is the total number of possible mappings,\footnote{It is
  $2N$ and not $N^2$ because, in polar coordinates, vertices are
  always connected to their two nearest neighbors in the $\theta$
  dimension.}and $\sigma_S^2=0.15$ is the variance of the
similarity. Example similarity curves are shown in
Figure~\ref{fig:model-traces}.

@@

To summarize, we approximate the likelihood term of
Equation~\ref{eq:lh-ratio} using the similarity function defined in
Equation~\ref{eq:similarity}. Because we assume mental rotations are
performed sequentially, this similarity can only be computed for the
actions listed in Equation~\ref{eq:actions}.

@@

In order to approximate Equation~\ref{eq:lh-ratio} using samples of
the similarity function, we must decide \textit{which} places to
sample and \textit{when} stop sampling. The models below differ in how
they make these decisions (see Figure~\ref{fig:model-traces} for
examples of their behavior).

@@

To determine the correct rotation, we solve for the rotation matrix by
computing $(\Xa \F_h^T)_\mathrm{left}^{-1}\cdot{}\Xb$, where
$(\Xa\F_h^T)_\mathrm{left}^{-1}$ is the left inverse of
$\Xa\F_h^T$. We then check each $h$ to see if the computation produces
a valid rotation matrix; the $h$ that does is the correct
hypothesis. This approach gives us the true value of $\theta$, so
Equation~\ref{eq:lh-ratio} becomes a generalized likelihood ratio
test, where $\theta$ is set to the MLE value, rather than being
marginalized:
\begin{equation}
  \ell = \frac{\max_\theta p(\Xb\ \vert\ \Xa, \theta, \hi)p(\hi)}{\max_\theta p(\Xb\ \vert\ \Xa, \theta, \hf)p(\hf)}.
  \label{eq:mle-lh-ratio}
\end{equation}

@@

What if the point correspondences are unknown? An intuitive and simple
model which obeys this constraint first rotate in the direction that
increases similarity, and then stop rotating once a ``match'' is
found, as determined by a threshold on the value of $S$. If all
rotations have been exhausted, then flip, and try rotating again. We
assume that the locations where $S$ is greater than the threshold
correspond to the true $\theta$ (or points near the true
$\theta$). So, as with the \Oc{} model, Equation~\ref{eq:lh-ratio}
becomes a likelihood ratio.

@@

In the current formulation of the problem, choosing the threshold is
straightforward because we know both the exact geometry of the shapes
and that a linear transformation exists which will perfectly align
them. However, this choice is not always clear \textit{a priori}, as
the global optimum depends on many factors (e.g., shape complexity,
dimensionality, perceptual uncertainty, and whether the shapes are
identical).  One way to deal with the problem of choosing a threshold
would simply be to perform an exhaustive search or use a global
optimization strategy. However, these strategies would not result in
the linear response time found by \citeA{Shepard1971}. A second
alternative, which could potentially demonstrate the linear response
time result, might instead just perform a Hill Climbing (\Hc{}) search
for a \textit{local} maximum. As with the \Oc{} and \Th{} models,
Equation~\ref{eq:lh-ratio} becomes a generalized likelihood ratio
test.

@@

While the previous few models have all focused on \textit{searching}
for the global maximum, the goal is really to just approximate
Equation~\ref{eq:lh-ratio} while still obeying the constraint of
sequential rotations. Based on evidence that people use ``active
sampling'' \cite<e.g.,>{Nelson2007,Gureckis:2012gu}, we hypothesize a
model that actively monitors the usefulness of the rotations it is
performing. Instead of searching for a maximum, we maintain a
probability distribution over our \textit{estimate} of
Equation~\ref{eq:lh-ratio}, and then sample actions which are expected
to improve that estimate.  This strategy has several benefits. First,
it does not make assumptions about the shape or scale of the
similarity function. Second, by choosing to sample places which are
informative, this method implicitly minimizes the amount of rotation.

Formally, we denote $Z_h$ as our estimate of the likelihood for
hypothesis $h$, and write its distribution as:
\begin{equation}
  p(Z_h) = \int \left[\int S(\Xb, f(\Xa, \theta, h))p(\theta)\dif\theta\right] p(S)\dif S,
\end{equation}
where $S$ is the similarity function, and $p(S)$ is a prior over
similarity functions.  This method of estimating an integral is known
in the machine-learning literature as \textit{Bayesian Quadrature}
\cite{Diaconis:1988uo,Osborne:2012tm}, or BQ.

@@

Assuming independence, we can now write $p(Z_h)\approx\mathcal{N}(Z_h\
\vert\ \mathbb{E}[Z_h], \mathbb{V}(Z_h))$, which gives us a
distribution over the likelihood ratio in Equation~\ref{eq:lh-ratio}:
\begin{equation}
p(\ell)\approx\frac{\mathcal{N}(Z_0\ \vert\ \mathbb{E}[Z_0], \mathbb{V}(Z_0))\ p(\hi)}{\mathcal{N}(Z_1\ \vert\ \mathbb{E}[Z_1], \mathbb{V}(Z_1))\ p(\hf)}.
\end{equation}
This distribution cannot easily be calculated; however, we really are
just interested in whether $Z_0>Z_1$ or $Z_1>Z_0$. Thus, rather than
computing $p(\ell)$, we use $Z_D=Z_0-Z_1$ and compute:
$p(Z_D)\propto\mathcal{N}(\mathbb{E}[Z_0] - \mathbb{E}[Z_1],
\mathbb{V}(Z_0) + \mathbb{V}(Z_1))$.

We then sample new observations until we are at least 95\% confident
that $Z_D\neq 0$. In other words, when $p(Z_D<0)<0.025$, we accept
$\hi$, and when $p(Z_D<0)>0.975$, we accept $\hf$. To determine which
points to sample, we can compute the expected variance of $Z_h$ if we
sampled a new observation at $\theta_a$. From \citeA{Osborne:2012tm}:
\begin{align}
\mathbb{E}&[\mathbb{V}(Z_h|\theta_a)]=\mathbb{V}(Z_h) + \mathbb{E}[Z_h] - \\
&\int \mathbb{E}[Z_h|\theta_{a}]^2 \mathcal{N}(\mu_h(\theta_a), \mathrm{Cov}_h(\theta_a, \theta_a))\dif\log S_h(\theta_a).\nonumber
\label{ref:expected-variance}
\end{align}
We compute this expected variance for each of the actions in
Equation~\ref{eq:actions}, and then choose the action with the lowest
value.

@@

We randomly generated 20 shapes of five or six vertices (e.g.,
Figure~\ref{fig:model-traces}). For each shape, we computed 20
``same'' and 20 ``flipped'' stimuli pairs, with 18 rotations
($\theta$) spaced at $20^\circ$ increments between $0^\circ$ and
$360^\circ$ (with $0^\circ$ and $180^\circ$ repeated twice, in order
to gather an equal number of responses for each angle between
$0^\circ$ and $180^\circ$). ``Same'' pairs were created by rotating
$\Xa$ by $\theta$; ``flipped'' pairs were first reflected $\Xa$ across
the $y$-axis, then rotated by $\theta$.

@@

\begin{figure*}[t]
  \begin{center}
    \includegraphics[width=\textwidth]{../../figures/D/response_time_scatters.pdf}
    \caption{\textbf{Model vs. human RTs.} Each subplot shows the
      z-scored model response times ($x$-axis) vs. the z-scored human
      response times ($y$-axis). Pearson correlations are shown
      beneath each subplot. The dotted lines are $x=y$.}
    \label{fig:human-model-scatters}
  \end{center}
\end{figure*}

@@

For each model, we ran 50 samples for each of the 800 experimental
stimuli. The step size parameter ($\sigma_\epsilon$) was fit to human
RTs for each of the models, resulting in $\sigma_\epsilon=0.6$ for the
\Th{} and \Bq{} models and $\sigma_\epsilon=0.1$ for the \Oc{} and
\Hc{} models. We also ran the models under two different priors,
$p(h=0)=0.5$ (the ``equal'' prior) and $p(h=0)=0.55$ (the ``unequal''
prior). This only had a major effect on the stopping criteria for the
\Bq{} model.\footnote{The prior also had a small effect on the
  accuracy of the \Hc{} model, because sometimes the local maxima were
  small enough that Equation~\ref{eq:mle-lh-ratio} became dominated by
  the prior.}

@@

For analyses of RT, confidence intervals around harmonic means were
computed using a bootstrap analysis using 10000 bootstrap samples
(sampled with replacement). RTs were only considered for correct
responses.  We also used a bootstrap analysis of 10000 bootstrap
samples to compute the confidence intervals around both Spearman
($\rho$) and Pearson ($r$) correlations.  Unless otherwise specified,
all correlations were computed over 720 datapoints (one for each
stimulus). For analyses of accuracy, confidence intervals were
computed from a binomial proportion with a Jeffrey's beta prior.  To
test if participants or a model was above chance on a particular
stimulus, we used the same binomial proportion, and simply tested
whether $p>\frac{0.05}{720}$ for the 50\% mark, where the denominator
is a Bonferroni correction for multiple comparisons.

@@

The average RT across all stimuli was \ExpTime{} milliseconds.  The
minimum angle of rotation was significantly rank-order (Spearman)
correlated with average per-stimulus RTs, both for ``flipped''
(\ExpThetaTimeCorrFlipped{}) and ``same'' pairs
(\ExpThetaTimeCorrSame{}). While this replicates the general result of
previous experiments \cite<e.g.,>{Shepard1971,Cooper:1975wp}, our
results are not as perfectly linear. However, \citeA{Gardony:2013gn}
found a constant effect for ``flipped'' pairs; an average of their
results and the classic results would produce something similar to
ours.

@@

To gauge between-participant reliability, we randomly split
participants into two groups, computed the average RT and accuracy per
stimulus for each group, and then computed correlations for these two
measures. We repeated this procedure 100 times, and from those samples
computed the median correlations and 95\% confidence intervals. For
RT, the between-participants correlation was \ExpTimeCorr{}, and for
accuracy, it was \ExpAccuracyCorr{}.

@@

As we would expect, the number of actions taken by the \Oc{} model was
perfectly correlated with the minimum angle of rotation
(Figure~\ref{fig:response-time-accuracy}). Unsurprisingly, the \Oc{}
model was the best fit to human RTs, with a correlation of
\OcTimeCorr{} (Figure~\ref{fig:human-model-scatters}).  However, the
\Oc{} model was 100\% accurate, and therefore the correlation between
minimum angle and accuracy was undefined.

@@

There was an overall monotonic relationship between the minimum angle
of rotation and the number of actions taken by the \Th{} model
(Figure~\ref{fig:response-time-accuracy}). However, the relationship
between angle and RT for individual shapes was usually not monotonic
(e.g., Figure~\ref{fig:response-time-stimulus}).  The \Th{} model was
able to explain a moderate amount of the variance in human RTs, with a
correlation of \ThTimeCorr{}
(Figure~\ref{fig:human-model-scatters}). Like the \Oc{} model, the
\Th{} model had 100\% accuracy, and thus there was no correlation
between minimum angle and accuracy.

As noted above, we fit $\sigma_\epsilon=0.6$ for the \Th{} model. This
has the interesting effect of actually causing the \Th{} to
\textit{over}rotate (see, for example, the ``flipped'' pair trace in
Figure~\ref{fig:model-traces}). This is because the step size is large
enough that it sometimes misses the global maximum, and must do
another full rotation to find it. However, when the step size is small
enough to always find the maximum, the fit to human data is slightly
worse.

@@

\begin{figure}[t]
  \begin{center}
    \includegraphics[width=0.48\textwidth]{../../figures/D/response_time_stimulus.pdf}
    \caption{\textbf{Typical RT curves for a single shape.}  These
      plots correspond to the shape shown in
      Figure~\ref{fig:model-traces}. Left: human curves are either
      linear (as with the ``same'' pairs), or linear and then flat (as
      with the ``flipped'' pairs). Middle: the \Th{} model does not
      have a reliably monotonic relationship with rotation. Right: the
      \Bq{} model is roughly linear.}
    \label{fig:response-time-stimulus}
  \end{center}
\end{figure}

The \Hc{} was the only model for which there was no monotonic
relationship between rotation and RT
(Figure~\ref{fig:response-time-accuracy}). Moreover, the \Hc{} model
was barely above chance (\HcAccuracy{}) and there were \HcNumChance{}
individual stimuli for which it was not above chance. The \Hc{} model
was not a particularly good predictor of human RTs (\HcTimeCorr{}), as
shown in Figure~\ref{fig:human-model-scatters}. It was a moderately
good predictor of human accuracy (\HcAccuracyCorr{}).

@@

Like the \Oc{} and \Th{} models, there was an overall monotonic
relationship between rotation and the number of steps taken by the
\Bq{} model (Figure~\ref{fig:response-time-accuracy}). Unlike the
\Th{} model, this relationship existed for individual shapes as well
(e.g., Figure~\ref{fig:response-time-stimulus}).  The \Bq{} model
explained variance in human RTs about as well as the \Th{} model
(Figure~\ref{fig:human-model-scatters}), with a correlation of
\BqTimeCorr{} for the equal prior and \BqpTimeCorr{} for the unequal
prior.

The \Bq{} model was quite accurate overall (equal prior:
(\BqAccuracy{}; unequal prior: \BqpAccuracy{}). With the equal prior,
there were \BqNumChance{} stimulii for which it was not above chance;
with the unequal prior, there were \BqpNumChance{}. The \Bq{} model
had about the same correlation with people's accuracy as the \Hc{}
model (equal prior: \BqAccuracyCorr{}; unequal prior:
\BqpAccuracyCorr{}).

Because the \Bq{} model relies on Equation~\ref{eq:lh-ratio} for its
stopping criteria (as opposed to just finding a maximum), the prior
$p(h)$ had a fairly significant effect, as shown in
Figure~\ref{fig:response-time-accuracy}. With just a small bias of
$p(h=0)=0.55$, there was a clear separation in RTs for ``same'' versus
``flipped'' stimuli; this separation is similar to the trend also
observed in human RTs. While the prior also had an effect on accuracy,
this effect did not reflect the trends in human behavior.

@@

The \Oc{} model was the best predictor of human RTs, but is an
unsatisfying explanation of human behavior because it relies on
\textit{a priori} knowledge that people are unlikely to have about
which points on the two shapes correspond. Similarly, while the \Th{}
was next best in terms of correlation, it makes the unsatisfying
assumption that the global maximum is known. Finally, neither of these
models can explain the difference in people's behavior on ``same'' and
``flipped'' stimuli.  The \Bq{} model was nearly as good as the \Th{}
model; in contrast to the other models, the \Bq{} model makes no
assumptions about people's \textit{a prior} knowledge. Furthermore,
the \Bq{} model's adaptive stopping rule is sensitive to the prior,
and thus provides a possible explanation for why people are slower to
respond on ``flipped'' stimulus pairs.

Additionally, a closer look at performance on individual shapes
reveals several interesting points
(Figure~\ref{fig:response-time-stimulus} shows curves for one
particular shape; these curves are fairly typical relative to other
shapes). First, the curves for the human RTs on specific shapes tend
to either be linear, or linear first, then flat. Second, while the
\Th{} model appears to capture the concavity of human RTs overall, it
is actually fairly inconsistent within shapes. Third, the linearity of
the \Bq{} model's RT curves \textit{is} consistent within shapes,
though it does not exhibit the same ``plateau'' effect that sometimes
appears in the human curves.

@@
